---
title: "Movie Recommendation System"
author: "Sourabh Teli"
date: "December 2, 2018"
output: html_document
---

```{r}
library(recommenderlab)
library(ggplot2)
library(data.table)
library(reshape2)
```
#Retrieving the Data
```{r}
setwd("C:/Users/Suraj/Desktop/IMDB Rating and project") 
movie_data <- read.csv("movies.csv",stringsAsFactors=FALSE)
rating_data <- read.csv("ratings.csv")
```
#Get to know about the Data
```{r}
summary(movie_data)
```
```{r}
summary(rating_data)
```
```{r}
head(movie_data)
```

```{r}
head(rating_data)
```
#Data Pre-Processing

Here from the summary we observe that the userId column and movieId column, consist of integers. Furthermore, we need to convert the genres present in the movie_data dataframe into a more usable format. In order to do so, we will first create a one-hot encoding to create a matrix that comprises of corresponding genres for each of the films.
```{r}
movie_genre <- as.data.frame(movie_data$genres, stringsAsFactors=FALSE)
library(data.table)
movie_genre2 <- as.data.frame(tstrsplit(movie_genre[,1], '[|]', 
                                   type.convert=TRUE), 
                         stringsAsFactors=FALSE) 
colnames(movie_genre2) <- c(1:10)
list_genre <- c("Action", "Adventure", "Animation", "Children", 
                "Comedy", "Crime","Documentary", "Drama", "Fantasy",
                "Film-Noir", "Horror", "Musical", "Mystery","Romance",
                "Sci-Fi", "Thriller", "War", "Western")
genre_mat1 <- matrix(0,10330,18)
genre_mat1[1,] <- list_genre
colnames(genre_mat1) <- list_genre
for (index in 1:nrow(movie_genre2)) {
  for (col in 1:ncol(movie_genre2)) {
    gen_col = which(genre_mat1[1,] == movie_genre2[index,col])
    genre_mat1[index+1,gen_col] <- 1
}
}
genre_mat2 <- as.data.frame(genre_mat1[-1,], stringsAsFactors=FALSE) #remove first row, which was the genre list
for (col in 1:ncol(genre_mat2)) {
  genre_mat2[,col] <- as.integer(genre_mat2[,col]) #convert from characters to integers
} 
str(genre_mat2)
```
we will create a search matrixâ that will allow us to perform an easy search of the films by specifying the genre present in our list.
```{r}
SearchMatrix <- cbind(movie_data[,1:2], genre_mat2[])
head(SearchMatrix)
```
Convert rating matrix into a recommenderlab sparse matrix to make sense of ratings for recommendation system. 
```{r}
ratingMatrix <- dcast(rating_data, userId~movieId, value.var = "rating", na.rm=FALSE)
ratingMatrix <- as.matrix(ratingMatrix[,-1]) #remove userIds
#Convert rating matrix into a recommenderlab sparse matrix
ratingMatrix <- as(ratingMatrix, "realRatingMatrix")
ratingMatrix
```
Let's check various options that are available to build recommendation system...
```{r}
recommendation_model <- recommenderRegistry$get_entries(dataType = "realRatingMatrix")
names(recommendation_model)
```
Let's check for the description of thease various options.
```{r}
lapply(recommendation_model, "[[", "description")
```
Here We will use Recommendations based on item-based collaborative filtering.

```{r}
recommendation_model$IBCF_realRatingMatrix$parameters
```
As Collaborative filtering mainly consider similarity between users to recommend new things, first we need to find similarities between preferences of various users.
check for percent similarities between first four users using similarity matrix to check for similarities. 
```{r}
similarity_mat <- similarity(ratingMatrix[1:4, ],
                               method = "cosine",
                               which = "users")
as.matrix(similarity_mat)
image(as.matrix(similarity_mat), main = "User's Similarities")

```
```{r}
movie_similarity <- similarity(ratingMatrix[, 1:4], method =
                                 "cosine", which = "items")
as.matrix(movie_similarity)
image(as.matrix(movie_similarity), main = "Movies similarity")
```
We need to consider or check for the ratings now..
```{r}
rating_values <- as.vector(ratingMatrix@data)
unique(rating_values)
```
```{r}
Table_of_Ratings <- table(rating_values) # creating a count of movie ratings
Table_of_Ratings
```
Now check for views of most watched movies...
```{r}
library(ggplot2)
movie_views <- colCounts(ratingMatrix) # count views for each movie
table_views <- data.frame(movie = names(movie_views),
                          views = movie_views) # create dataframe of views
table_views <- table_views[order(table_views$views,
                                 decreasing = TRUE), ] # sort by number of views
table_views$title <- NA
for (index in 1:10325){
  table_views[index,3] <- as.character(subset(movie_data,
                                         movie_data$movieId == table_views[index,1])$title)
}
table_views[1:6,]
```
Now check for movies with most views... 
```{r}
ggplot(table_views[1:5, ], aes(x = title, y = views)) +
  geom_bar(stat="identity", fill = 'steelblue') +
  geom_text(aes(label=views), vjust=-0.1, size=5.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("Total Views of the Top Films")
```
```{r}
image(ratingMatrix[1:20, 1:25], axes = FALSE, main = "Heatmap of the first 25 rows and 25 columns")
```

#Data Preparation 
We will conduct data preparation in the following three steps -

Selecting useful data.
Normalizing data.
Binarizing the data.

For finding useful data in our dataset, we have set the threshold for the minimum number of users who have rated a film as 50. This is also same for minimum number of views that are per film. This way, we have filtered a list of watched films from least-watched ones.

```{r}
movie_ratings <- ratingMatrix[rowCounts(ratingMatrix) > 50,
                             colCounts(ratingMatrix) > 50]
movie_ratings
```
From the above output of 'movie_ratings', we observe that there are 420 users and 447 films as opposed to the previous 668 users and 10325 films. We can now delineate our matrix of relevant users as follows -

```{r}
minimum_movies<- quantile(rowCounts(movie_ratings), 0.98)
minimum_users <- quantile(colCounts(movie_ratings), 0.98)
image(movie_ratings[rowCounts(movie_ratings) > minimum_movies,
                     colCounts(movie_ratings) > minimum_users],
main = "Heatmap of the top users and movies")
```

```{r}
average_ratings <- rowMeans(movie_ratings)
qplot(average_ratings, fill=I("steelblue"), col=I("red")) +
  ggtitle("Distribution of the average rating per user")
```
#Data Normalization 
Some users give high ratings and some give low ratings,this can create bias while building model..so we normalize the ratings data
```{r}
normalized_ratings <- normalize(movie_ratings)
sum(rowMeans(normalized_ratings) > 0.00001)
image(normalized_ratings[rowCounts(normalized_ratings) > minimum_movies,
                          colCounts(normalized_ratings) > minimum_users],
main = "Normalized Ratings of the Top Users")
```
#Item Based Collaborative Filtering System 

we will split our data as training and testing data

```{r}
sampled_data<- sample(x = c(TRUE, FALSE),
                      size = nrow(movie_ratings),
                      replace = TRUE,
                      prob = c(0.8, 0.2))
training_data <- movie_ratings[sampled_data, ]
testing_data <- movie_ratings[!sampled_data, ]
```
Now we will check for various parameters of our item based collaborative filter...
```{r}
recommendation_system <- recommenderRegistry$get_entries(dataType ="realRatingMatrix")
recommendation_system$IBCF_realRatingMatrix$parameters
```
#Model
```{r}
recommen_model <- Recommender(data = training_data,
                          method = "IBCF",
                          parameter = list(k = 30))
recommen_model
class(recommen_model)
```
Now we will retrive the model and check for class and dimension of our similarity matrix within our model...
```{r}
model_info <- getModel(recommen_model)
class(model_info$sim)
dim(model_info$sim)
top_items <- 20
image(model_info$sim[1:top_items, 1:top_items],
   main = "Heatmap of the first rows and columns")
```
Now we will carry out the sum of rows and columns with the similarity of the objects above 0. We will visualize the sum of columns through a distribution as follows -
```{r}
sum_rows <- rowSums(model_info$sim > 0)
table(sum_rows)
sum_cols <- colSums(model_info$sim > 0)
qplot(sum_cols, fill=I("steelblue"), col=I("red"))+ ggtitle("Distribution of the column count")
```
#Recommendation System
Here is a model with predict() function  that will identify similar items and will rank them appropriately. Here, each rating is used as a weight. Each weight is multiplied with related similarities. Finally, everything is added in the end.
```{r}
top_recommendations <- 5 # the number of items to recommend to each user
predicted_recommendations <- predict(object = recommen_model,
                          newdata = testing_data,
                          n = top_recommendations)
predicted_recommendations
```
```{r}
user1 <- predicted_recommendations@items[[1]] # recommendation for the first user
movies_user1 <- predicted_recommendations@itemLabels[user1]
movies_user2 <- movies_user1
for (index in 1:5){
  movies_user2[index] <- as.character(subset(movie_data,
                                         movie_data$movieId == movies_user1[index])$title)
}
movies_user2
```


